DRAFT: Harmonized and trait aggregation paper

### 2020-10-28 Paper/Draft_harmonization_paper.tex
25.10: Minor updates
28.10: Changes before send to Ralf

In this study, we follow the terminology proposed by Schmera et al., where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: Description of affinity
% A trait can described using a binary coding (exists or not), or using affinity scores that indicate the which extent an organism expresses a certain trait.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To our knowledge, only Brown et al. \cite{brown_functional_2018} harmonized grouping features from more than two geographically distant invertebrate trait databases in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. 
% TODO: ?Paper trait convergence Doledec

We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required \cite{schneider_towards_2019}. In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonized using standardized terminology. However, a lack of standardized terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology \cite{baird_toward_2011, schneider_towards_2019}. Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternative from Ben:
Traits can be described in a binary fashion with no uncertainty how the trait is expressed in any particularly organism (e.g. adult terrestrial stage, presence of gills), continuous (e.g. tolerance of pollution, body size) or using multiple descriptions (e.g. feeding mode) with some uncertainty as to which trait the organism should be coded to. One approach for dealing with the uncertainty in the latter is the use of fuzzy coded variables. Fuzzy codes are used to account for plasticity in traits, variability in traits within taxonomic groups above species, and incomplete knowledge and are usually converted to proportions.

@@

Traits of the harmonized grouping feature datasets were aggregated to family-level using three approaches. I) Direct aggregation of taxa to family-level giving equal weight to every taxon using the mean or median, denoted \textit{direct\_agg\textsubscript{mean}} and \textit{direct\_agg\textsubscript{median}}, respectively. II) Stepwise aggregation, i.e. first to the genus-level and subsequently to the family-level using the mean or median. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{mean}} or \textit{stepwise\_agg\textsubscript{median}}, respectively. III) Aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights the genera according to the number of their species in the trait datasets regardless if for every used grouping feature information was available.

@@

\begin{figure}
  \centering
  \subfile{Flowchart/Flowchart_methods.tex}
  \caption{Data processing steps of the selected traits. Intermediate (gray) and main (orange) steps of data preparation are depicted. The dashed bottom box illustrates the different trait aggregation methods using a small made-up example (data in the upper left corner). The aggregation methods (blue) and intermediate steps of the aggregation methods (purple) are displayed. Abbreviations: EU: Europe, NOA: North America, AUS: Australia, NZ: New Zealand, DB: Database.}
\end{figure}

@@

\begin{itemize}
  \item \textbf{aggregated vs assigned traits}
  \begin{itemize}
    \item In general, differences produced by the aggregation methods seem to arise by using mean or median, not so much driven by the different weighting approaches (direct vs. stepwise vs. weighted)
    \item Approaches that used the median resulted in fewer differences to the assigned traits, albeit if differences occurred to the assigned traits, they were greater. 
    \item Simulation shows that approaches using the median generally produce a larger range of values $\rightarrow$ Especially with higher trait variability for situations where one genus has a lot more species than the other genera within the family 
    \item Differences greater for the NOA dataset because of the binary coding $\rightarrow$ differences are greater between traits
    \item Taxonomic hierarchy in the investigated datasets? \\
     $\rightarrow$ In every investigated dataset, most families have only one genus  
    % -> Similarly, many genera have only a few species
    % -> linear relationship: the more genus per family, the more species (although there are exceptions -> e.g. Europe)
    \item There is no "true" value to compare to, i.e. assigned traits can be biased. Comparison of species-lvl traits in fwe and traits on genus-lvl in tachet would be interesting in this context? 
    \item  Suggestions from Co-Authors: \\
     $\rightarrow$ should include where the classifications come from \\
     $\rightarrow$ what do we really know because someone has watched the animals and published their behaviour; what is extrapolated knowledge? \\
     $\rightarrow$ are the differences between continents because of evolution or only because scientists have done more research in one of the continents?
  \end{itemize}
  \item \textbf{Influence of aggregation methods on trait-environment relationships}
  \begin{itemize}
    \item We re-analysed only a small-scale study, what would be different when applying a large-scale study?
  \end{itemize}
\end{itemize}

>>>>

In this study, we follow the terminology proposed by Schmera et al., where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". 
% TODO: Description of affinity

To our knowledge, only Brown et al. \cite{brown_functional_2018} harmonized grouping features from more than two geographically distant invertebrate trait databases in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates.

We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required \cite{schneider_towards_2019}. In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonized using standardized terminology. However, a lack of standardized terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology \cite{baird_toward_2011, schneider_towards_2019}. Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases.

% Alternative from Ben:
Traits can be described in a binary fashion with no uncertainty how the trait is expressed in any particularly organism (e.g. adult terrestrial stage, presence of gills), continuous (e.g. tolerance of pollution, body size) or using multiple descriptions (e.g. feeding mode) with some uncertainty as to which trait the organism should be coded to. One approach for dealing with the uncertainty in the latter is the use of fuzzy coded variables. Fuzzy codes are used to account for plasticity in traits, variability in traits within taxonomic groups above species, and incomplete knowledge and are usually converted to proportions.

@@

Traits of the harmonized grouping feature datasets were aggregated to family-level using three approaches. I) Direct aggregation of taxa to family-level giving equal weight to every taxon using the mean or median, denoted \textit{direct\_agg\textsubscript{mean}} and \textit{direct\_agg\textsubscript{median}}, respectively. II) Stepwise aggregation, i.e. first to the genus-level and subsequently to the family-level using the mean or median. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{mean}} or \textit{stepwise\_agg\textsubscript{median}}, respectively. III) Aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights the genera according to the number of their species in the trait datasets regardless if for every used grouping feature information was available (Figure \ref{fig:data_proc_overview}).

@@

\begin{figure}
  \centering
  \subfile{Flowchart/Flowchart_methods.tex}
  \caption{Data processing steps of the selected traits. Intermediate (gray) and main (orange) steps of data preparation are depicted. The dashed bottom box illustrates the different trait aggregation methods using a small made-up example (data in the upper left corner). The aggregation methods (blue) and intermediate steps of the aggregation methods (purple) are displayed. Abbreviations: EU: Europe, NOA: North America, AUS: Australia, NZ: New Zealand, DB: Database.}
  \label{fig:data_proc_overview}
\end{figure}

@@

\begin{itemize}
  \item \textbf{aggregated vs assigned traits}
  \begin{itemize}
    \item In general, differences produced by the aggregation methods seem to arise by using mean or median, not so much driven by the different weighting approaches (direct vs. stepwise vs. weighted)
    \item Approaches that used the median resulted in fewer differences to the assigned traits, albeit if differences occurred to the assigned traits, they were greater. 
    \item Simulation shows that approaches using the median generally produce a larger range of values $\rightarrow$ Especially with higher trait variability for situations where one genus has a lot more species than the other genera within the family 
    \item Differences greater for the NOA dataset because of the binary coding $\rightarrow$ differences are greater between traits
    \item Taxonomic hierarchy in the investigated datasets? \\
     $\rightarrow$ In every investigated dataset families have only 1 to few genera. 76 (North America) to 93 \% (New Zealand) of the families have 5 or less genera (see SI). 
    \item There is no "true" value to compare to, i.e. assigned traits at genus or family level can be biased as well (i.e. the person who assigned the trait was thinking about particular species within a genus not all). Comparison of species-level traits in fwe and traits on genus-level in tachet would be interesting in this context? 
    \item  Suggestions from Co-Authors: \\
     $\rightarrow$ should include where the classifications come from \\
     $\rightarrow$ what do we really know because someone has watched the animals and published their behaviour; what is extrapolated knowledge? \\
     $\rightarrow$ are the differences between continents because of evolution or only because scientists have done more research in one of the continents?
  \end{itemize}
  \item \textbf{Influence of aggregation methods on trait-environment relationships}
  \begin{itemize}
    \item We re-analysed only a small-scale study, what would be different when applying a large-scale study?
  \end{itemize}
\end{itemize}
